### 2023-06-13 python_excel_ite.tex
working on first notebook

@@

At our institution, this module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics. Basic Linux skills are also part of the PCDA course. The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.  The EwP module could certainly be used as part of a semester long Python based analytics course. The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed. We will describe the AAP course web in more detail below in the context of the EwP module.

What kind of Python background do students need to have in order to be ready for the EwP module? This module presumes a basic familiarity with Python fundamentals for data analysis such as:

\begin{itemize}
	\item variables, artithmetic and boolean operators, and basic data types,
	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
	\item flow control such as branching with \texttt{if \ldots else} and looping with \texttt{for} and \texttt{while}
	\item module imports,
	\item using built in functions and accessing methods and properties of objects,
	\item creating functions,
	\item basic use of NumPy and pandas for data wrangling and analysis,
	\item basic plotting with matplotlib.
\end{itemize}

@@

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages. GIVE URL. Some background on Jupyter notebooks for analytics. Questions and coding challenges sprinkled throughout the notebooks - answers at bottom. While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course. The screencasts are highly valued by the students as they can rewatch things as needed. Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.

@@

There are two overarching learning objectives for this first set of Jupyter notebooks. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis. At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and then starting over and doing things in a different way. Each notebook is accompanied by a few to several videos that walk the students through the material - a guided tour with challenges to overcome along the way. The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.

@@

@@

The first focuses on building a simple what-if model and doing sensitivity analysis much like one would do using Excel's Data-Table functionality.

This example is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/). In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation. Here is the basic problem (with a few modifications):

@@

The notebook begins with a suggestion to start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, we store base input values in variables.

@@

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, we can rely on the built in vectorized behavior of the language. A range of order quantities is created as a NumPy array and then can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate (loop) through a collection of input values.

>>>>

Integrated or separate lit review section?

@@

At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken my ``Practical Computing for Business Analytics'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5470}{PCDA}) course in which they spend about half the semester learning R and the other half learning Python, both in the context of business analytics. Basic Linux skills are also part of the PCDA course. The PCDA course ends with intro to the sklearn library for doing predictive modeling in Python and the AAP course begins where the PCDA course ends.  The EwP module could certainly be used as part of a semester long Python based analytics course. The students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have extensive course websites that are publicly accessible, including access to all of the video content and files needed. We will describe the AAP course web in more detail below in the context of the EwP module.

What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is needed, including:

\begin{itemize}
	\item variables, arithmetic and boolean operators, and basic data types,
	\item data structures such as lists, dictionaries, tuples, NumPy arrays and pandas dataframes,
	\item flow control such as branching with \texttt{if \ldots else} and looping with \texttt{for} and \texttt{while}
	\item module imports,
	\item using built in functions and accessing methods and properties of objects,
	\item creating functions,
	\item basic use of NumPy and pandas for data wrangling and analysis,
	\item basic plotting with matplotlib.
\end{itemize}

@@

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks, all publicly accessible from extensive course web pages. GIVE URL. Some background on Jupyter notebooks for analytics - see downloaded papers. Questions and coding challenges sprinkled throughout the notebooks - answers at bottom. While AAP is online asynch, this same approach has been used very successfully in a computer teaching lab for the PCDA course. The screencasts are highly valued by the students as they can rewatch things as needed. Many are working professionals and a 2.5 hour evening class in a warm computer lab can be taxing.

@@

There are two overarching learning objectives for this first set of Jupyter notebooks. Students learn how Python might be used to do something they are quite familiar with doing in a spreadsheet - model building and sensitivity analysis. At the same time, students encounter, as they are needed, more advanced Python programming techniques such as object oriented programming and manipulation of tuples, lists and dictionaries. In this way, the computational problem to be solved drives the introduction of more advanced Python concepts and techniques instead of such things being presented in a vacuum. I make it a point to try to convey the notion that there are often multiple possible approaches by embarking on one path and there can be value to starting over and doing things in a different way. Each notebook is accompanied by videos that walk the students through the material and act as a guided tour with challenges to overcome along the way. The screenshot in \ref{fig:activities} shows the portion of the course web page that provides the high level roadmap through the notebooks and associated screencasts.

@@

The stage is set with an overview of a typical spreadsheet based model.

@@

\begin{figure}[!ht]
	\centering
	\begin{minipage}{.80\textwidth}
		\includegraphics[width=1.0\textwidth]{images/output_2_0.png}
	\end{minipage}
	\caption{Typical spreadsheet model}
	\label{fig:ldr_pvsa}
\end{figure}

The example used throughout the rest of the notebook is based on one in the [spreadsheet modeling textbooks I've used in my classes since 2001](https://host.kelley.iu.edu/albrightbooks/). In both books, they introduce the "Walton Bookstore" problem in the chapter on Monte-Carlo simulation. Here is the basic problem (with a few modifications):

@@

We start with a very simple model that ignores the uncertainty in demand and that is not object oriented. Proceeding much like we would in Excel, we store base input values in variables.

@@

Since order quantity is the key decision variable, we might want to see how profit changes for different order quantities in this first simplified model. In Excel, the Data Table tool provides an easy way to do this and is a staple of spreadsheet based management science textbooks. In a Data Table, a range of order quantities can be used as row or column input and one or more output values can be computed. In Python, we can rely on the vectorized behavior of the language. A range of order quantities is created as a NumPy array that can be used directly to compute vectors of all of the intermediate and final output variables. Much like Data Tables, vectorized computations avoid having to explicitly iterate, or loop, through a collection of input values.
